Contribution Title\thanks{Supported by organization x.}

### 2019-10-24 paper/paper.tex
Save the hboxes!

\begin{minipage}{.34\textwidth}
\begin{lstlisting}[language={[objective]caml},numbers=left,basicstyle=\small\tt]
match 1 with
  | 1 -> match "one" with
         | str -> str
  | 2 -> "two"
\end{lstlisting}
\end{minipage}
\vrule
\begin{minipage}{\textwidth}
\begin{lstlisting}[basicstyle=\small\tt]
 File "./nestmatch.ml", line 4, characters 4-5:
 Error: This pattern matches values of type int
        but a pattern was expected which
        matches values of type string
\end{lstlisting}
\end{minipage}

@@

\noindent This is a resolvable ambiguity, since $\parse($\verb|'1 + (2 + 3)'|$) = \{t_1\}$ and $\parse($\verb|'(1 + 2) + 3'|$) = \{t_2\}$. To demonstrate the unresolvable case, we add the production $E -> s: E$ \verb|';'| $E$, at which point we find that the word \verb|'[1 ; 2]'| is unresolvably ambiguous; $\parse($\verb|'[1 ; 2]'|$) = \{t_3, t_4\}$ where:

@@

\begin{itemize}
\item $s$ and $f$ are two new distinct states.
\item $Q = \{s, f\} \cup \bigsqcup_l Q_l$.
\item $\T' = \T \cup \{'\pospl', '\pospr'\}$.
\item $\Gamma = Q \times Q \times \Labels^{*}$.
\item $\delta$ is defined by the following equations (unspecified cases produce $\emptyset$):
  $$
  \begin{array}{r@{\;=\;}l@{\;\mid\;}l}
    \delta(p, a, \lambda)
    & \multicolumn{2}{@{}l@{}}{\{(\delta_l(p, a), \lambda)\} \qquad \text{ where } p \in Q_l \text{ and } a \in \T} \\

    \delta(s, \text{``}\pospl\text{''}, \lambda)
    & \{ (s_{l}, (s, f, w))
    & (S, w, N) \in T, N = \labelnt(l) \} \\

    \delta(p, \text{``}\pospl\text{''}, \lambda)
    & \{ (s_{l_3}, (p, q, w))
    & p \in Q_{l_1}, \delta_{l_1}(p, N) = q, (N, w, N') \in T, N' = \labelnt(l_3) \} \\

    \delta(p, \text{``}\pospr\text{''}, (s, f, w))
    & \{ (f, \lambda)
    & (S, w, N) \in T, N = \labelnt(l), p \in F_l \} \\

    \delta(q', \text{``}\pospr\text{''}, (p, q, w))
    & \{ (q, \lambda)
    & q' \in F_{l_1}, \labelnt(l_1) = N', (N, w, N') \in T, p \in Q_{l_2}, \delta_{l_2}(p, N) = q \} \\

@@

\begin{tabular}{ll}
\small
\begin{lstlisting}[language=syncon,boxpos=t]
type Exp
grouping "(" Exp ")"
precedence {
  mul; // higher in precedence list
  add; // means higher precedence
}
\end{lstlisting}
&
\small
\begin{lstlisting}[language=syncon,boxpos=t]
token Integer = "[0-9]+"
syncon literal: Exp = n:Integer
syncon list: Exp =
  "[" (head:Exp (";" tail:Exp)*)? "]"
infix add: Exp = "+"
infix mul: Exp = "*"
\end{lstlisting}
\end{tabular}\smallskip

@@

\begin{itemize}
\item We could decide to forbid the case where two comparisons are
  right next to each other, e.g., \syncon{forbid gt.left = gt} and
  \syncon{forbid gt.right = gt} (where \syncon{gt} is the syncon
  for the greater-than operator).
\item We could change the definition of the parallel combinators
  and define separate lexical tokens for the combinators, e.g.,
  \syncon{token Seq = ">[[:lower:]][[:word:]]*>"}. This would add
  just enough whitespace sensitivity to allow separating the
  different cases.
\item We could change the syntax of the combinators to avoid
  clashes with other operators.
\end{itemize}

@@

\paragraph{Current Limitations of Syncons}
%
Other than syntactic OCaml constructs that are not yet specified,
there are two sources of failure in the parts of the OCaml
compiler that we cannot parse.
%
First, our parsing tool does not support specifying ``longest
match'' on a production, which is required to handle the pattern
matching constructs correctly. Some cases can be handled using
\syncon{forbid} declarations, but it does not get us all the way.
%
Second, our system uses a different definition of precedence than
the OCaml language. Our translation from precedence to implicit
\syncon{forbid} declarations is shallow (it only considers direct
children), while the OCaml has deep precedence. For example,
addition binds stronger than \ocaml{let} (which syntactically
functions as a prefix operator in OCaml), and thus
%
``\ocaml{1 + let x = 1 in x + 2}'' should be parsed as
%
``\ocaml{1 + (let x = 1 in (x + 2))}''. Our tool, however, reports
the expression as ambiguous, additionally suggesting the
alternative interpretation
%
``\ocaml{(1 + (let x = 1 in x)) + 2}''.
%
This interpretation is indeed valid if we only look at the direct
children of each operator, since ``\ocaml{(_ + _) + 2}'',
``\ocaml{1 + let _ in _}'', and ``\ocaml{let x = 1 in x}'' are all
individually correct with regards to precedence.

@@

\begin{itemize}
\item Visibly pushdown languages with the same input partitions
  are closed under intersection, complement, and union
  \cite{alurVisiblyPushdownLanguages2004}. Intersection is given
  by a product automaton. Given a pair of VPDAs
  $(Q_1, \T, \delta_1, q_0, F_1)$ and
  $(Q_2, \T, \delta_2, q'_0, F_2)$ their product automaton has the
  form $(Q_1 \times Q_2, \T, \delta', (q_0, q'_0),$
  $ F_1 \times F_2)$ where, when $c \in \T_c$, $i \in \T_i$ and
  $r \in \T_r$:
  \[
  \begin{array}{r@{\ =\ }l}
    \delta'((p_1, p_2), c, \lambda) & \{((q_1, q_2), (g_1, g_2)) \mid (q_1, g_1) \in \delta_1(p_1, c, \lambda), (q_2, g_2) \in \delta_2(p_2, c, \lambda) \}\\
    \delta'((p_1, p_2), i, \lambda) & \{((q_1, q_2), \lambda) \mid (q_1, \lambda) \in \delta_1(p_1, i, \lambda), (q_2, \lambda) \in \delta_2(p_2, i, \lambda) \}\\
    \delta'((p_1, p_2), r, (g_1, g_2)) & \{((q_1, q_2), \lambda) \mid (q_1, \lambda) \in \delta_1(p_1, r, \lambda), (q_2, \lambda) \in \delta_2(p_2, r, \lambda) \}\\
  \end{array}
  \]

>>>>

\begin{minipage}{.34\textwidth}
\begin{lstlisting}[language={[objective]caml},numbers=left,basicstyle=\small\tt]
match 1 with
  | 1 -> match "one" with
         | str -> str
  | 2 -> "two"
\end{lstlisting}
\end{minipage}
\vrule
\begin{minipage}{\textwidth}
\begin{lstlisting}[basicstyle=\small\tt]
 File "./nest.ml", line 4, characters 4-5:
 Error: This pattern matches values of type
        int but a pattern was expected which
        matches values of type string
\end{lstlisting}
\end{minipage}

@@

\noindent This is a resolvable ambiguity, since $\parse($\verb|'1 + (2 + 3)'|$) = \{t_1\}$ and \\$\parse($\verb|'(1 + 2) + 3'|$) = \{t_2\}$. To demonstrate the unresolvable case, we add the production $E -> s: E$ \verb|';'| $E$, at which point we find that the word \verb|'[1 ; 2]'| is unresolvably ambiguous; $\parse($\verb|'[1 ; 2]'|$) = \{t_3, t_4\}$ where:

@@

\begin{itemize}
\item $s$ and $f$ are two new distinct states.
\item $Q = \{s, f\} \cup \bigsqcup_l Q_l$.
\item $\T' = \T \cup \{'\pospl', '\pospr'\}$.
\item $\Gamma = Q \times Q \times \Labels^{*}$.
\item $\delta$ is defined by the following equations (unspecified cases produce $\emptyset$):
  % TODO: This could be made prettier and more aligned
  $$
  \begin{array}{l}
    \delta(p, a, \lambda) = \{(\delta_l(p, a), \lambda)\} \qquad \text{ where } p \in Q_l \text{ and } a \in \T \\

    \delta(s, \text{``}\pospl\text{''}, \lambda) =
    \{ (s_{l}, (s, f, w)) \mid
    (S, w, N) \in T, N = \labelnt(l) \} \\

    \delta(p, \text{``}\pospl\text{''}, \lambda) = \\
    \qquad\{ (s_{l_3}, (p, q, w)) \mid
    p \in Q_{l_1}, \delta_{l_1}(p, N) = q, (N, w, N') \in T, N' = \labelnt(l_3) \} \\

    \delta(p, \text{``}\pospr\text{''}, (s, f, w)) =
    \{ (f, \lambda) \mid
    (S, w, N) \in T, N = \labelnt(l), p \in F_l \} \\

    \delta(q', \text{``}\pospr\text{''}, (p, q, w)) = \\
    \qquad \{ (q, \lambda) \mid
    q' \in F_{l_1}, \labelnt(l_1) = N', (N, w, N') \in T, p \in Q_{l_2}, \delta_{l_2}(p, N) = q \} \\

@@

\begin{tabular}{ll}
\small
\begin{lstlisting}[language=syncon,boxpos=t]
type Exp
grouping "(" Exp ")"
precedence {
  mul; // higher in list means
  add; // higher precedence
}
\end{lstlisting}
&
\small
\begin{lstlisting}[language=syncon,boxpos=t]
token Integer = "[0-9]+"
syncon literal: Exp = n:Integer
syncon list: Exp =
  "[" (head:Exp (";" tail:Exp)*)? "]"
infix add: Exp = "+"
infix mul: Exp = "*"
\end{lstlisting}
\end{tabular}\smallskip

@@

\begin{itemize}
\item We could decide to forbid the case where two comparisons are
  right next to each other, e.g., \syncon{forbid gt.left = gt} and
  \syncon{forbid gt.right = gt} (where \syncon{gt} is the syncon
  for the greater-than operator).
\item We could change the definition of the parallel combinators
  and define separate lexical tokens for the combinators, e.g,
  \syncon{token Seq = ">[[:lower:]]}\\\syncon{[[:word:]]*>"}. This
  would add just enough whitespace sensitivity to allow separating
  the different cases.
\item We could change the syntax of the combinators to avoid
  clashes with other operators.
\end{itemize}

@@

\paragraph{Current Limitations of Syncons}
%
Other than syntactic OCaml constructs that are not yet specified,
there are two sources of failure in the parts of the OCaml
compiler that we cannot parse.
%
First, our parsing tool does not support specifying ``longest
match'' on a production, which is required to handle the pattern
matching constructs correctly. Some cases can be handled using
\syncon{forbid} declarations, but it does not get us all the way.
%
Second, our system uses a different definition of precedence than
the OCaml language. Our translation from precedence to implicit
\syncon{forbid} declarations is shallow (it only considers direct
children), while the OCaml has deep precedence. For example,
addition binds stronger than \ocaml{let} (which syntactically
functions as a prefix operator in OCaml), and thus
%
``\ocaml{1 + let x = 1 in x + 2}'' should be parsed as
%
``\ocaml{1 + (let x = 1 in (x + 2))}''. Our tool, however, reports
the expression as ambiguous, additionally suggesting the
alternative interpretation
%
``\ocaml{(1 + (let x}\\\ocaml{= 1 in x)) + 2}''.
%
This interpretation is indeed valid if we only look at the direct
children of each operator, since ``\ocaml{(_ + _) + 2}'',
``\ocaml{1 + let _ in _}'', and ``\ocaml{let x = 1 in x}'' are all
individually correct with regards to precedence.

@@

\begin{itemize}
\item Visibly pushdown languages with the same input partitions
  are closed under intersection, complement, and union
  \cite{alurVisiblyPushdownLanguages2004}. Intersection is given
  by a product automaton. Given a pair of VPDAs
  $(Q_1, \T, \delta_1, q_0, F_1)$ and
  $(Q_2, \T, \delta_2, q'_0, F_2)$ their product automaton has the
  form $(Q_1 \times Q_2, \T, \delta', (q_0, q'_0),$
  $ F_1 \times F_2)$ where, when $c \in \T_c$, $i \in \T_i$ and
  $r \in \T_r$: \RED{TODO: Is this definition correct? $g_1$ and
    $g_2$ are not used in the third case}
  \[
  \begin{array}{l}
    \delta'((p_1, p_2), c, \lambda) =\\ \qquad\{((q_1, q_2), (g_1, g_2)) \mid (q_1, g_1) \in \delta_1(p_1, c, \lambda), (q_2, g_2) \in \delta_2(p_2, c, \lambda) \}\\
    \delta'((p_1, p_2), i, \lambda) =\\ \qquad \{((q_1, q_2), \lambda) \mid (q_1, \lambda) \in \delta_1(p_1, i, \lambda), (q_2, \lambda) \in \delta_2(p_2, i, \lambda) \}\\
    \delta'((p_1, p_2), r, (g_1, g_2)) =\\ \qquad \{((q_1, q_2), \lambda) \mid (q_1, \lambda) \in \delta_1(p_1, r, \lambda), (q_2, \lambda) \in \delta_2(p_2, r, \lambda) \}\\
  \end{array}
  \]
